Macro Impacts

### 2016-07-23 Macro Impacts Revised.tex
Concept of Program

\pagebreak
\section{Numerical Methods}
From symmetry, it is clear that it is enough to understand the behavior of rays from a point source to understand the propagation of rays in the whole sphere. Let $\vec p=(r,\theta)$ denote the position vector of our ray, whose initial coordinates are $\vec p(0)=(x_0,0)$. We also insist that the tangent vector of $\vec p$ be at an angle $\alpha$ from the x axis. The parametrization in terms of these initial conditions is
\begin{align}
r(s)&=\f{a}{b}\sqrt{\f{a^2\p{e^{2abs}-1}^2+b^2\p{e^{2abs}+1}^2x_0^2+2(-1)^{s_1}ab\p{e^{4abs}-1}x_0\abs{\cos\alpha}}{a^2\p{e^{2abs}+1}^2+b^2\p{e^{2abs}-1}^2x_0^2+2(-1)^{s_1}ab\p{e^{4abs}-1}x_0\abs{\cos\alpha}}}\,,\\
\notag\theta(s)&=\arccot\ps{\f{4(-1)^{s_2}abx_0(a^2-b^2x_0^2)\sin\alpha}{4a^2b^2(1+e^{4abs})x_0^2+4(-1)^{s_1}abe^{4abs}x_0(a^2+b^2x_0^2)\cos\alpha+(e^{4abs}-1)(a^4+b^4x_0^4+2a^2b^2x_0^2\cos(2\alpha))}}-\\
&\hspace{1cm}-\arctan\ps{\f{(-1)^{s_2}(2abx_0+(-1)^{s_1}(a^2+b^2x_0^2)\cos\alpha)\csc\alpha}{a^2-b^2x_0^2}}
\end{align}
Note that $s_1$ and $s_2$ change the quadrant of the ray propagation with respect to the starting position. We now need to solve for $s$ such that $r$ is some desired value at that point, call it $r_m$ (where we have chosen $m$ to stand for mantle).
\\\\
The time $s_m$ it takes to get to the radius $r_m$ is
\begin{align}
s_m=\f{1}{2ab}\ln\ps{\f{a^4-b^4 r_m^2 x_0^2+a^2b^2(r_m^2-x_0^2)+(-1)^{s_3} 2ab\sqrt{(r_m^2-x_0^2)(a^4-b^4 r_m^2x_0^2)+(a^2-b^2r_m^2)^2x_0^2\cos^2\alpha}}{(a^2-b^2r_m^2)(a^2+b^2x_0^2+2(-1)^{s_1}abx_0\cos\alpha)}}
\end{align}
We now want to compute the sine of the angle between the tangent vector to the curve and the surface normal at $s_m$. Observe
\begin{align}
T\cdot n&=\f{\dot r}{\sqrt{\dot r^2+r^2\dot\theta^2}}\,.
\end{align}
Hence
\begin{align}
\sin\phi&=\f{r\dot\theta}{\sqrt{\dot r^2+r^2\dot\theta^2}}\,.
\end{align}
We now write a theoretical program for computing the energy density on the surface of the earth. First, define a list of $M$ points from which we want to propagate rays. In our specific case, these will be evenly spaced points along a chord through the earth. Store the coordinates of these points. From each of these points, we will want to propagate some number $N$ of rays. These rays must have uniform angular separation. This is the question of tiling the sphere by arbitrarily small polygons. This is, of course, only possible for finitely many polygons on a given sphere. Thus, we need some other way of uniformly distributing points on a sphere. {http://neilsloane.com/packings/dim3/pack.3.130.txt} contains the coordinates of 130 points placed on a sphere such that their angular separation is minimized (18.5 degrees). If this isn't enough, there is a tiling on the same website with over 1000 points, but that would be absurd. The vectors given by the 130 points determine the plane of propagation (given by the span of the direction vector and the vector pointing from the origin) and the initial propagation angle $\alpha$ with respect to the vector pointing from the origin.

>>>>

\section{Numerical Methods}
From symmetry, it is clear that it is enough to understand the behavior of rays from a point source to understand the propagation of rays in the whole sphere. Let $\vec p=(r,\theta)$ denote the position vector of our ray, whose initial coordinates are $\vec p(0)=(x_0,0)$. We also insist that the tangent vector of $\vec p$ be at an angle $\alpha$ from the x axis. The parametrization in terms of these initial conditions is
r(s)&=\f{a}{b}\sqrt{\f{a^2\p{e^{2abs}-1}^2+b^2\p{e^{2abs}+1}^2x_0^2+2(-1)^{s_1}ab\p{e^{4abs}-1}x_0\abs{\cos\alpha}}{a^2\p{e^{2abs}+1}^2+b^2\p{e^{2abs}-1}^2x_0^2+2(-1)^{s_1}ab\p{e^{4abs}-1}x_0\abs{\cos\alpha}}}\,,\\
\notag\theta(s)&=\arccot\ps{\f{4(-1)^{s_2}abx_0(a^2-b^2x_0^2)\sin\alpha}{4a^2b^2(1+e^{4abs})x_0^2+4(-1)^{s_1}abe^{4abs}x_0(a^2+b^2x_0^2)\cos\alpha+(e^{4abs}-1)(a^4+b^4x_0^4+2a^2b^2x_0^2\cos(2\alpha))}}-\\
&\hspace{1cm}-\arctan\ps{\f{(-1)^{s_2}(2abx_0+(-1)^{s_1}(a^2+b^2x_0^2)\cos\alpha)\csc\alpha}{a^2-b^2x_0^2}}
Note that $s_1$ and $s_2$ change the quadrant of the ray propagation with respect to the starting position. We now need to solve for $s$ such that $r$ is some desired value at that point, call it $r_m$ (where we have chosen $m$ to stand for mantle).
The time $s_m$ it takes to get to the radius $r_m$ is
s_m=\f{1}{2ab}\ln\ps{\f{a^4-b^4 r_m^2 x_0^2+a^2b^2(r_m^2-x_0^2)+(-1)^{s_3} 2ab\sqrt{(r_m^2-x_0^2)(a^4-b^4 r_m^2x_0^2)+(a^2-b^2r_m^2)^2x_0^2\cos^2\alpha}}{(a^2-b^2r_m^2)(a^2+b^2x_0^2+2(-1)^{s_1}abx_0\cos\alpha)}}
We now want to compute the sine of the angle between the tangent vector to the curve and the surface normal at $s_m$. Observe
T\cdot n&=\f{\dot r}{\sqrt{\dot r^2+r^2\dot\theta^2}}\,.
Hence
\sin\phi&=\f{r\dot\theta}{\sqrt{\dot r^2+r^2\dot\theta^2}}\,.
We now write a theoretical program for computing the energy density on the surface of the earth. First, define a list of $M$ points from which we want to propagate rays. In our specific case, these will be evenly spaced points along a chord through the earth. Store the coordinates of these points. From each of these points, we will want to propagate some number $N$ of rays. These rays must have uniform angular separation. This is the question of tiling the sphere by arbitrarily small polygons. This is, of course, only possible for finitely many polygons on a given sphere. Thus, we need some other way of uniformly distributing points on a sphere. {http://neilsloane.com/packings/dim3/pack.3.130.txt} contains the coordinates of 130 points placed on a sphere such that their angular separation is minimized (18.5 degrees). If this isn't enough, there is a tiling on the same website with over 1000 points, but that would be absurd. The vectors given by the 130 points determine the plane of propagation (given by the span of the direction vector and the vector pointing from the origin) and the initial propagation angle $\alpha$ with respect to the vector pointing from the origin. Once we have the set of points and rays to consider, our problem turns into many simpler problems, i.e. rays propagating in two dimensions. There are four cases to consider.
\begin{enumerate}
\item $r(0)>r_m$
\begin{enumerate}
\item $s_m\in\C\setminus\R$. For these, it is enough to know $\theta(s_E)$, where $s_E$ is the time at which the rays reach the surface. In the case that we want to consider surface reflection, we should store the angle of incidence, but that is likely unnecessary for our purposes.
\item $s_m\in\R$. These rays will reflect off the core. The angle of reflection with respect to the surface normal is simply minus the angle of incidence. Store these data and re-propagate ray after all first-iteration propagation is done.
\end{enumerate}
\item $r(0)<r_m$
\begin{enumerate}
\item If the angle of incidence is less than the critical angle, then the ray refracts. Store these data and re-propagate after all first iteration propagation is done.
\item If the angle of incidence is greater than the critical angle, the rays will internally reflect forever, so we can throw these out.
\end{enumerate}
\end{enumerate}
The first round of propagation will produce some rays that hit the surface and some which need to be propagated a second time. All those rays propagating in the second round of propagation will originate on the core-mantle boundary, and thus computation will be greatly simplified. Once we have stored the final polar coordinate, we must place these final coordinates properly on the sphere. This is done by two rotations. The first is the inverse of the rotation which took the initial propagation direction vector into the $xy$-plane. The second is the inverse of the rotation which took the position vector of the point source to the $x$-axis. I believe rotations of points is commutative, but I'm too lazy to check, so it's safest to just do these in the proper order.
